Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2018-08-14 paper/paper.tex
Update paper
Use latex template and add impreovements from meeting.

% !TeX document-id = {5ae73978-d7a8-4a15-a272-3b0597ff2241}
% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.19 (04-Sep-2017)

% !TeX spellcheck = en-US
% !TeX encoding = utf8
% !TeX program = pdflatex
% !BIB program = bibtex
% -*- coding:utf-8 mod:LaTeX -*-

%"runningheads" enables:
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
%This is good for other readers to enable proper archiving among other papers and pointing to
%content. Even if the title page states the title, when printed and stored in a folder, when
%blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
%it is good to have title printed on the pages, too.
\documentclass[runningheads,a4paper]{llncs}[2017/09/04]

%% If you need packages for other papers,
%% START COPYING HERE

% Set English as language and allow to write hyphenated"=words
%
% Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
% english has to go last to set it as default language
\usepackage[ngerman,english]{babel}
%Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}

\usepackage{ifluatex}
\ifluatex
  \usepackage{fontspec}
  \usepackage[english]{selnolig}
\fi

\iftrue % use default-font
  \ifluatex
    % use the better (sharper, ...) Latin Modern variant of Computer Modern
    \setmainfont{Latin Modern Roman}
    \setsansfont{Latin Modern Sans}
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % better font, similar to the default springer font
    % cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
    \usepackage[%
      rm={oldstyle=false,proportional=true},%
      sf={oldstyle=false,proportional=true},%
      tt={oldstyle=false,proportional=true,variable=false},%
      qt=false%
    ]{cfr-lm}
  \fi
\else
  % In case more space is needed, it is accepted to use Times New Roman
  \ifluatex
    \setmainfont{TeX Gyre Termes}
    \setsansfont[Scale=.9]{TeX Gyre Heros}
    % newtxtt looks good with times, but no equivalent for lualatex found,
    % therefore tried to replace with inconsolata.
    % However, inconsolata does not look good in the context of LNCS ...
    %\setmonofont[StylisticSet={1,3},Scale=.9]{inconsolata}
    % ... thus, we use the good old Latin Modern Mono font for source code.
    \setmonofont{Latin Modern Mono} % "variable=false"
    %\setmonofont{Latin Modern Mono Prop} % "variable=true"
  \else
    % overwrite cmodern with the Times variant
    \usepackage{newtxtext}
    \usepackage{newtxmath}
    \usepackage[zerostyle=b,scaled=.9]{newtxtt}
  \fi
\fi

\ifluatex
\else
  % fontenc and inputenc are not required when using lualatex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc} %support umlauts in the input
\fi

@@

% backticks (`) are rendered as such in verbatim environment. See https://tex.stackexchange.com/a/341057/9075 for details.
\usepackage{upquote}

% Nicer tables (\toprule, \midrule, \bottomrule - see example)
\usepackage{booktabs}

%extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

%put figures inside a text
%\usepackage{picins}
%use
%\piccaptioninside
%\piccaption{...}
%\parpic[r]{\includegraphics ...}
%Text...

% For easy quotations: \enquote{text}
% This package is very smart when nesting is applied, otherwise textcmds (see below) provides a shorter command
\usepackage{csquotes}

% For even easier quotations: \qq{text}
\usepackage{textcmds}

%enable margin kerning
\RequirePackage[%
  babel,%
  final,%
  expansion=alltext,%
  protrusion=alltext-nott]{microtype}%
% \texttt{test -- test} keeps the "--" as "--" (and does not convert it to an en dash)
\DisableLigatures{encoding = T1, family = tt* }

%tweak \url{...}
\usepackage{url}
%\urlstyle{same}
%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
%DO NOT ACTIVATE -> prevents line breaks
%\makeatletter
%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
%\makeatother

% Diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
% Slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
%\usepackage{diagbox}

% Required for package pdfcomment later
\usepackage{xcolor}

% For listings
\usepackage{listings}
\lstset{%
  basicstyle=\ttfamily,%
  columns=fixed,%
  basewidth=.5em,%
  xleftmargin=0.5cm,%
  captionpos=b}%
\renewcommand{\lstlistingname}{List.}
% Fix counter as described at https://tex.stackexchange.com/a/28334/9075
\usepackage{chngcntr}
\AtBeginDocument{\counterwithout{lstlisting}{section}}

% Enable nice comments
\usepackage{pdfcomment}
%
\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}
%
% Compatibality with packages todo, easy-todo, todonotes
\newcommand{\todo}[1]{\commentatside{#1}}
% Compatiblity with package fixmetodonotes
\newcommand{\TODO}[1]{\commentatside{#1}}

% Bibliopgraphy enhancements
%  - enable \cite[prenote][]{ref}
%  - enable \cite{ref1,ref2}
% Alternative: \usepackage{cite}, which enables \cite{ref1, ref2} only (otherwise: Error message: "White space in argument")

% Doc: http://texdoc.net/natbib
\usepackage[%
  square,        % for square brackets
  comma,         % use commas as separators
  numbers,       % for numerical citations;
%  sort,          % orders multiple citations into the sequence in which they appear in the list of references;
  sort&compress, % as sort but in addition multiple numerical citations
                 % are compressed if possible (as 3-6, 15);
]{natbib}
% In the bibliography, references have to be formatted as 1., 2., ... not [1], [2], ...
\renewcommand{\bibnumfmt}[1]{#1.}

\ifluatex
  % does not work when using luatex
  % see: https://tex.stackexchange.com/q/419288/9075
\else
  % Prepare more space-saving rendering of the bibliography
  % Source: https://tex.stackexchange.com/a/280936/9075
  \SetExpansion
  [ context = sloppy,
    stretch = 30,
    shrink = 60,
    step = 5 ]
  { encoding = {OT1,T1,TS1} }
  { }
\fi

% Put footnotes below floats
% Source: https://tex.stackexchange.com/a/32993/9075
\usepackage{stfloats}
\fnbelowfloat

% Enable that parameters of \cref{}, \ref{}, \cite{}, ... are linked so that a reader can click on the number an jump to the target in the document
\usepackage{hyperref}
% Enable hyperref without colors and without bookmarks
\hypersetup{hidelinks,
  colorlinks=true,
  allcolors=black,
  pdfstartview=Fit,
  breaklinks=true}
%
% Enable correct jumping to figures when referencing
\usepackage[all]{hypcap}

\usepackage[group-four-digits,per-mode=fraction]{siunitx}

%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}
\crefname{listing}{\lstlistingname}{\lstlistingname}
\Crefname{listing}{Listing}{Listings}

%Intermediate solution for hyperlinked refs. See https://tex.stackexchange.com/q/132420/9075 for more information.
\newcommand{\Vlabel}[1]{\label[line]{#1}\hypertarget{#1}{}}
\newcommand{\lref}[1]{\hyperlink{#1}{\FancyVerbLineautorefname~\ref*{#1}}}

\usepackage{xspace}
%\newcommand{\eg}{e.\,g.\xspace}
%\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.,\ }
\newcommand{\ie}{i.\,e.,\ }

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
  <-6>    MnSymbolC5
  <6-7>   MnSymbolC6
  <7-8>   MnSymbolC7
  <8-9>   MnSymbolC8
  <9-10>  MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

\ifluatex
\else
  % Enable copy and paste - also of numbers
  % This has to be done instead of \usepackage{cmap}, because it does not work together with cfr-lm.
  % See: https://tex.stackexchange.com/a/430599/9075
  \input glyphtounicode
  \pdfgentounicode=1
\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%% END COPYING HERE

% Add copyright
% Do that for the final version or if you send it to colleagues
\iffalse
  %state: intended|submitted|llncs
  %you can add "crop" if the paper should be cropped to the format Springer is publishing
  \usepackage[intended]{llncsconf}

  \conference{name of the conference}

  %in case of "llncs" (final version!)
  %example: llncs{Anonymous et al. (eds). \emph{Proceedings of the International Conference on \LaTeX-Hacks}, LNCS~42. Some Publisher, 2016.}{0042}
  \llncs{book editors and title}{0042} %% 0042 is the start page
\fi

% For demonstration purposes only
\usepackage[math]{blindtext}
\usepackage{mwe}
\usepackage{pgfplots}

\begin{document}

\title{Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics}
%If Title is too long, use \titlerunning
\titlerunning{Modeling and Simulation of Load Balancing Strategies}

%Single insitute
%\author{ \and Firstname Lastname}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
%\institute{Institute}

% Multiple insitutes - ALTERNATIVE to the above
 \author{
Ren\'e Caspart\inst{1} \and
Patrick Firnkes\inst{2} \and
Manuel Giffels\inst{1} \and
Anne Koziolek\inst{2} \and
G\"unter Quast\inst{1} \and
Ralf Reussner\inst{2} \and
Maximilian Stemmer-Grabow\inst{2}
 }

%
%If there are too many authors, use \authorrunning
\authorrunning{R. Caspart et al.}
%
\institute{
	Institute for Experimental Particle Physics (ETP) \and
	Institute for Program Structures and Data Organization (IPD) \\
	at Karlsruhe Institute of Technology (KIT)
}

\maketitle

\begin{abstract}
The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years. To cope with this increasing load, most efficient usage of the resources is mandatory. Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system. 
A first step to meet both challenges, a more efficient utilization of the grid and handling the rising complexity of the system, a simulation for the WLCG should be created. There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.

@@

\begin{keywords}
  Performance Prediction,
  Modeling,
  Palladio,
  HEP,
  Load Balancing Strategies
\end{keywords}

@@

The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the community of HEP \cite{community}.
\Cref{resources} shows the estimated CPU resources for the ATLAS experiment in the next ten years. As can be seen the flat computing budget model does not cover future needs for computing resource in the period from 2026 to 2028 due the start of the HL-LHC era. One of the key aspects to deal with this increasing load is to use the existing resources most efficiently.

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user job. As cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be used to compute them. This results in a more complex scheduling, because each resource has a different optimal job types and mixtures. Furthermore, this leads to a more dynamic data placement, where fewer centers hosts data and thus the remote access is increasing.

@@

To solve both challenges, a more efficient usage of the grid and handling the rising complexity of the system, a simulation of the WLCG should be created.
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulation does not obtain the necessary detail level to simulate effects on working nodes that required to evaluate load balancing strategies.

Optimizing load strategies without the help of simulations requires stable load, in order to preserve the comparability of the different strategies. This is neither possible to achieve for the WLCG nor for a single center like GridKa, because the nature of the load is highly dynamical. Another option would be to use an isolated test system for the evaluation. However, the system under inspection is too large and thus a similar test systems would not be financially affordable. A smaller test system could not deliver correct results because too much of the complexity would be taken away.
These reasons lead to the conclusion that a model and simulation of the performance of computing jobs is required to evaluate different load balancing strategies.

Our long term goal is to evaluate different load balancing strategies for the WLCG and additionally evaluate different design decisions affecting the performance of the grid. These gained insights will help the HEP community to solve the challenges caused by the HL-LHC upgrade.
As a proof of concept we modeled and simulated the performance of CMS computing jobs at the Tier 1 center GridKa.
The model and simulation will be created using the Palladio simulator. We extended this simulator to be able to simulate load balancing strategies. Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
This model enables us to evaluate strategies for GridKa and allows us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

This paper is organized as follows: \Cref{sec:palladio} presents the Palladio simulator as a foundation of the simulation. \Cref{sec:process} gives a overview on the simulation process.\Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio. \Cref{sec:param} explains how we determine the model parameters and the automated model creation. The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work. In \Cref{outlook} possible design decisions and how they could be modeled and simulated are presented. Finally, \Cref{conlusion} gives a conclusion on this paper.

@@

These models are the component model, the assembly model, the resource model, the allocation model, and the usage model.
The component model specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
To represent how the components are connected to model the software architecture, the assembly model has to be created. 
The resource model describes the resource environment. It contains the number and characteristics of the resource containers on which the components could be deployed and the topology of the network connecting these containers.
Allocation models represent the mapping of the different components to the resource containers.
Finally, the usage model defines how a system is used regarding workload, user behavior, and parameters.
Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.

Lehrig and Becker \cite{arch} extended Palladio with Architectural Templates to make it possible to model cloud computing environments efficiently.

Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
It was later extended to support the metrics scalability, elasticity, and efficiency \cite{arch}.

@@

\section{Simulation Process}
\label{sec:process}
The starting point of the simulation process is the analysis of the job and working node data to extract performance parameters for the model. For example, these are the CPU demand of jobs, the frequency of job types or the CPU power of the nodes. This analysis is automated and described in \Cref{sec:param}. The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters. Both steps can be done via the UI of Palladio and do not require manually modeling. At this point the users have a valid and calibrated model of GridKa with its resource and the workload. The properties of this model are described in \Cref{sec:model}.

@@

@@

After a job is scheduled on a resource container it tries to acquire the specified number of job slots or waits if the node has not sufficient free slots. After that the CPU and I/O demand of the job are processed. To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations, which represent the number of events. Thus, for each event there are I/O and CPU demands resources needed.
While processing, the job uses the specified number of cores to process the CPU demand. When the job finishes, it releases the job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

Also each type of working node in GridKa is modeled in the resource model. Each node has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type. These types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.

At GridKa all computing jobs can run on all nodes, thus each job component is deployed on each computing node.

The usage model defines which and how many jobs are scheduled. To make sure that there is always a high load on the system, we used a closed workload with 5000 jobs. Each job type has a configurable share of the overall load.

@@

The simulated throughput was higher than the measured one, because of that the we added a delay after each job, in order that the measured throughput matches the simulated one. The first factor for the higher throughput could be the missing overhead of pilot jobs. The second, more crucial, factor is the missing of jobs in the job data. We observed that 10-20\% of jobs are missing in the data, the cause for this is not known yet, however this should be investigated. 
After adding the delay the simulated throughput matches the measured one accurately, as \Cref{throughput} shows.
\Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.
\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/durchsatz}
	\caption[]{Measured and Simulated Throughput}
	\label{throughput}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/shares}
	\caption[]{Measured and Simulated Shares of Job Types}
	\label{shares}
\end{figure}

The simulated CPU efficiency is lower than the measured one, as shown in \Cref{utilization}. The reason for this is likely missing jobs in the job data. Because of that is the calibrated throughput is too low resulting in a lower utilization.

@@

\section{Related Work}
\label{related}
The MONARC project was started 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
However, after they initial design was developed, the project was discontinued in 2000. 
In 2004 a second version of the MONARC simulator was released with several improvements, such as simulation of data replication. However, MONARC2 showed scalability issues when simulating a Tier-2 site and accuracy issues regarding network load \cite{1742-6596-331-7-072038}.

There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
However, both of them are missing graphical model editors which makes the models of the computing jobs are time-consuming to create \cite{simgrid}\cite{cloudsim}. The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency.

@@

Korenkov et. al simulated the JINR Tier1 site in 2016 \cite{jinr}. They focused on the effects of the location of the required job data the throughput of the site. However, they neither modeled each working node nor modeled details like job slots, thus this simulation would not be usable to simulate load balancing strategies on working node level.

Hushchyn et al. created a simulation for the LHCb Grid in 2017 using SimGrid \cite{lhcb}. This simulation focused on which site a job should be placed based on data location. They modeled each site coarse granularly, thus it is not possible to simulate scheduling effects inside a computing center. In contrast our model is more detailed and is able to simulate effects regarding scheduling on working node level.

\section{Outlook}
\label{outlook}
The next steps are the evaluation of different design decisions using our simulation. In this section a few design decisions are described that could be simulated.

@@

Additionally, the effects of integrating cloud resources could be simulated. The model must be extended with the cloud resources and the required network load to load the files from a site. Then a good scheduling strategy for integration of cloud resources could be evaluated, like only scheduling CPU heavy jobs to cloud resources and keep I/O heavy jobs near the data.

Also the effects of data caches could be simulated. Firstly, several sites must be implemented to simulate their interactions. Secondly, the jobs must be parametrized with the possibility of remote reading. Then the data cache should be modeled. Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.
\section{Conclusion}
\label{conlusion}
We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is. Regarding scalability we are confident that several computing centers could be simulated in such a detailed manner like we did with the CMS computing jobs at GridKa. For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.

This proof of concept showed that an automated model parameter extraction and model creation is possible and that Palladio can be used to evaluate load balancing strategies for computing in HEP. In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level. Furthermore, we are confident that our approach helps to improve utilization of the grid. The next step is the evaluation of a design decision mentioned in \Cref{outlook}.

\renewcommand{\bibsection}{\section*{References}} % requried for natbib to have "References" printed and as section*, not chapter*
% Use natbib compatbile splncsnat style.
% It does provide all features of splncs03, but is developed in a clean way.
% Source: http://phaseportrait.blogspot.de/2011/02/natbib-compatible-bibtex-style-bst-file.html
\bibliographystyle{splncsnat}
\begingroup
  \ifluatex
    %try to activate if bibliography looks ugly
    %\sloppy
  \else
    \microtypecontext{expansion=sloppy}
  \fi
  \small % ensure correct font size for the bibliography
  \bibliography{paper}
\endgroup

% Enfore empty line after bibliography
\ \\
%

>>>>

\documentclass[a4paper]{jpconf}

@@

\newcommand{\TODO}[1]{\commentatside{#1}}
\usepackage[capitalise,nameinlink]{cleveref}
%Nice formats for \cref
\crefname{section}{Sect.}{Sect.}
\Crefname{section}{Section}{Sections}
\crefname{listing}{\lstlistingname}{\lstlistingname}
\Crefname{listing}{Listing}{Listings}
\usepackage{pgfplots}

\bibliographystyle{iopart-num}

\begin{document}
\title{Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics}

\author{Ren\'e Caspart${}^{1}$,
	Patrick Firnkes${}^{2}$,
	Manuel Giffels${}^{1}$,
	Anne Koziolek${}^{2}$,
	G\"unter Quast${}^{1}$,
	Ralf Reussner${}^{2}$ and 
	Maximilian Stemmer-Grabow${}^{2}$}

\address{
	${}^{1}$Institute for Experimental Particle Physics (ETP) \\
	${}^{2}$Institute for Program Structures and Data Organization (IPD) \\
	at Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany}

\ead{koziolek@kit.edu}

\begin{abstract}
	The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years. To cope with this increasing load, most efficient usage of the resources is mandatory. Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system. 
	A first step to meet both challenges, a more efficient utilization of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created. There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.

@@

@@

The amount of data processed by experiments in high energy physics (HEP) will drastically increase in the coming years due the upgrade of the LHC to the High-Luminosity Large Hadron Collider (HL-LHC), creating major challenges for the community of HEP \cite{community}.
\Cref{resources} shows the estimated required and provided CPU resources for the ATLAS experiment in the next ten years. Assuming a flat computing budget model, the estimated provided resources do not cover future needs for computing resource in the period from 2026 to 2028 due the start of the HL-LHC era. One of the key aspects to deal with this increasing load is the more efficient usage of existing resources.

Another challenge for the HEP community is the increasing heterogeneity and distribution of the computing resources of user jobs. Cloud resources, like HNSciCloud and AWS, or institute clusters and Tier 3 centers will be progressively introduced for the computation of these jobs. This results in a more complex scheduling, because each resource has a different optimal job types and mixtures. Furthermore, this leads to a more dynamic data placement, where fewer centers hosts data and thus the remote access is increasing.

@@

To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system, a simulation of the WLCG should be created.
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.

Optimizing load balancing strategies without the help of simulations requires stable load, in order to preserve the comparability of the different strategies. This is neither possible to achieve for the WLCG nor for a single center like GridKa, because the nature of the load is highly dynamical. Another option would be to use an isolated test system for the evaluation. However, the system under inspection is too large and thus a similar test systems would not be financially affordable. A smaller test system could not deliver correct results because too much of the complexity of the system would be taken away.
These reasons lead to the conclusion that a model and simulation of the performance of computing jobs is required to evaluate different load balancing strategies.

Our long term goal is to evaluate different load balancing strategies for the WLCG and additionally evaluate different design decisions affecting the performance of the grid. These gained insights will help the HEP community to solve the challenges caused by the HL-LHC upgrade.

The contribution of this paper is the modeling and simulating the performance of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator. We extended this simulator to be able to simulate load balancing strategies. Furthermore, we created a automated model parameter analysis and model creation, to increase the usability.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

This paper is organized as follows: \Cref{sec:palladio} presents the Palladio simulator as a foundation of the simulation. \Cref{sec:process} gives a overview on the simulation process. \Cref{sec:model} describes how different elements of the computing jobs and the computing resources are represented in our model and the implementation of the necessarily extension of Palladio. \Cref{sec:param} explains how we determine the model parameters and the automated model creation. The validation is presented in \Cref{validation} and \Cref{related} puts our work in context with the related work. In \Cref{outlook} possible design decisions and how they could be modeled and simulated are presented. Finally, \Cref{conlusion} gives a conclusion on this paper.

@@

These models are the component model, the assembly model, the resource model, the allocation model, and the usage model.
\begin{itemize}
	\item The component model: Specifies the structure and behavior of the components independently from their later usage. This allows reuse of the components, but requires its parametrization.
	\item The assembly model: Represents how the components are connected to model the software architecture.
	\item The resource model: Describes the resource environment. It contains the number and characteristics of the resource containers on which the components could be deployed and the topology of the network connecting these containers.
	\item The allocation model: Represents the mapping of the different components to the resource containers.
	\item The usage model: Defines the usage of a system regarding workload, user behavior, and parameters.

\end{itemize}
Once these models are created, Palladio can be used to simulate the system choosing one of several supported simulators.

Lehrig and Becker \cite{arch} extended Palladio with Architectural Templates to make it possible to model cloud computing environments efficiently.
Furthermore, Palladio's simulation approach SimuLizar was created to simulate self-adaptive systems \cite{becker2013simulizar}.
It was later extended to support the metrics scalability, elasticity, and efficiency \cite{arch}.

@@

\section{Simulation Process}
\label{sec:process}

\Cref{process} shows the simulation process starting with the matching of the data sets to receiving a valid model of the site and its jobs.

The starting point of the simulation process is to match the data sets from different data sources as a prerequisite  of the analysis. In our proof of concept we have three data sources: The CERN WMArchive, CMS GridKa job data and GridKa node performance data.

After matching and grouping the data for job types the analysis of the job and working node data to extract performance parameters for the model can take place. For example, these are the CPU demand of jobs, the frequency of the job types or the CPU power of the nodes. This analysis and the data set matching are automated and described in \Cref{sec:param}.

The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters. Both steps can be done via the UI of Palladio and do not require manually modeling. At this point the users have a valid and calibrated model of GridKa with its resource and the workload. The properties of this model are described in \Cref{sec:model}.

@@

\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/process}
	\caption[]{The Simulation Process}
	\label{process}
\end{figure}

@@

After a job is scheduled on a resource container it acquires the specified number of job slots or waits if the node has not sufficient free slots. After that the CPU and I/O demand of the job are processed. To avoid one big I/O operation at the start of a job and make the processing of I/O more continuously, the processing is modeled using iterations that reflect the number of events. Thus, for each iteration there are I/O and CPU resources needed. Rather than simulating each event, we simulate a part of the processed events, because the simulation of each event would harm the scalability of the simulation. 
While processing, the job uses the specified number of cores to process the CPU demand. To simulate not explicitly modeled elements of the jobs such as the overhead created by pilot jobs we added a delay to each job. This delay is calibrated in order to match the measured throughput. After the job finished, it releases its acquired job slots. All of these parameters are specified as a probability function and are extracted from measured performance data as described in \Cref{sec:param}.

Each type of working node in GridKa is modeled in the resource model. Each node type has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type. The types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.

At GridKa every type of computing job can run on every node, thus each job type is allocated on each computing node in the allocation model.

The usage model defines which and how many jobs are scheduled. To make sure that there is always a high load on the system, we used a closed workload with enough jobs that the system is fully utilized. Each job type has a configurable share of the overall load.

@@

As mentioned in \Cref{sec:model} a delay was added to the job to simulate not explicitly modeled elements. This delay was calibrated so that the measured throughput matches the simulated one of the calibration dataset. The simulated throughput of the test dataset matches the measured one accurately with only 1\% difference.
\Cref{shares} depicts the share of the simulated and measured job types, which also have a high accuracy.

\begin{figure}
	\centering
	\includegraphics[scale = 0.4]{images/shares}
	\caption[]{Measured and Simulated Shares of Job Types}
	\label{shares}
\end{figure}

The simulated CPU efficiency is lower than the measured one, as shown in \Cref{utilization}. The reason for this are likely missing jobs in the job data. We observed that a significant amount of data is missing, the cause for this is not known yet, however this should be investigated. Because of that the calibrated throughput is too low, resulting in a lower utilization. Another reason is the missing utilization created by pilot jobs that are not explicitly modeled yet.

@@

\section{Related Work}
\label{related}
The MONARC project was started 1998 to help with the development of the initial design of the WLCG at CERN \cite{monarc2000models}.
However, after the initial design was developed, the project was discontinued in 2000 and thus no model and simulation of the WLCG was created. 
In 2004 a second version of the MONARC simulator was released with several improvements, such as simulation of data replication. However, MONARC2 showed scalability issues when simulating a Tier-2 site and accuracy issues regarding network load \cite{1742-6596-331-7-072038}.

There are a variety of grid and cloud simulators existing, of which SimGrid and CloudSim are two popular ones. 
However, both of them are missing graphical model editors which makes the modeling of the computing jobs time-consuming \cite{simgrid}\cite{cloudsim}. The support of graphical editors is a big advantage of Palladio and increases its usability and efficiency. 
 parameterize

@@

Korenkov et. al simulated the JINR Tier1 site in 2016 \cite{jinr} and Hushchyn et al. created a simulation for the LHCb Grid in 2017 using SimGrid \cite{lhcb}. Both simulations focused on the effects of the location of the required job data on the throughput of the sites.

However, they neither modeled each working node nor modeled details of the scheduling process like job slots, thus both simulations would not be usable to simulate load balancing strategies on working node level.

\section{Outlook}
\label{outlook}
Our next steps are the evaluation of different design decisions. In this section a few design decisions are described that could be simulated.

@@

Additionally, the effects of integrating cloud resources could be simulated. The model must be extended with the cloud resources and the required network load to access the files from a site. Then a good load balancing strategy for the  integration of cloud resources could be evaluated, like only scheduling CPU heavy jobs to cloud resources and keeping I/O heavy jobs near the data.

Also the effects of data caches could be simulated. Firstly, several sites must be modeled to simulate their interactions. Secondly, the jobs must be parametrized with the possibility of remote access. Then the data cache should be modeled. Finally, the performance improvements of the data cache by transferring data from the cache and not from sites can be simulated.
\section{Conclusion}
\label{conlusion}
We successfully simulated the CMS computing jobs at GridKa. To enable easy usage of the simulation we implemented an automated model parameter extraction and model creation. Furthermore, we extended Palladio to be able to simulate load balancing strategies. Our first results suggest that this approach could deliver accurate results regarding CPU efficiency, however it must be investigated what the reason for the data inconsistency and missing job information is. Regarding scalability we are confident that several computing centers could be simulated on a detailed level. For use cases that require the simulation of larger parts of the WLCG we suggest a more macroscopic model that focuses on the interaction of the sites and models the sites internals on a less detailed level.

This proof of concept showed that an automated model parameter extraction and model creation is possible and that Palladio can be used to evaluate load balancing strategies for computing in HEP. In contrast to the existing simulations our simulation models a site in a more detailed way, enabling the evaluation of load balancing strategies on working node level. Furthermore, we are confident that our approach helps to improve utilization of the grid and thus helps solving the challenges created by the HL-LHC update. The next step is the evaluation of a design decision mentioned in \Cref{outlook}.

\section*{References}
\bibliography{paper}
